Region-network sparsity priors for high-dimensional inference in fMRI data

### 2015-11-25 paper/paper.tex
intro outline standing

\title{Introducing region-network priors to statistical estimators for fMRI data}

@@

\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
  Ga\"el Varoquaux, Bertrand Thirion
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

@@

SPECIAL
Functional specialization has been investigated in many ways.
%
Single cell recordings and microscopic examination
revealed, for instance, the
specialization in the visual cortex into V1, V2, V3, V3A, and V4
\citep{hubel1962receptive, zeki1978functional}.

example: fusiform gyrus
-> lesion -> does not work anymore

Non-invasive brain imaging with
functional magnetic resonance imaging (fMRI) and
positron emission tomography (PET)
enabled the localization of
sensory, motor, and emotional functions to cortical areas
in the living brain
\citep{fristen1997imaging}.

Radioactive mapping the distribution of
neurotransmitter receptors added yet another
local characteristic of neuronal populations
\citep{zilles2009receptor}.

clustering
\citep{behrens03}

High-thoughput computing enabled an 
ultahigh-resolution 3D model of brain anatomy
at macroscopical to near-cellular scale
\citep{amunts2013bigbrain}.

=> methods: non-overlapping, discrete region units

@@

@@

@@

>>>>

\title{Region-network sparsity priors for high-dimensional inference in fMRI data}

@@

\author{Danilo Bzdok, Karl Zilles, Michael Eickenberg,
  Ga\"el Varoquaux, Bertrand Thirion
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

@@

SPECIAL
Functional specialization has been
explored with many different research methods.
%
Single cell recordings and microscopic examination
revealed, for instance, the
specialization in the visual cortex into V1, V2, V3, V3A, and V4
\citep{hubel1962receptive, zeki1978functional}.
Tissue lesion reports of the mid-fusiform gyrus of the visual system
frequently indicated impaired
recognition of others' identity from faces
\citep{iaria2008contribution}.
The localization of
sensory, motor, and emotional cognitive functions to cortical areas
in the living brain
has later been enabled by
non-invasive brain imaging with
functional magnetic resonance imaging (fMRI) and
positron emission tomography (PET)
\citep{fristen1997imaging}.
Further,
radioactive mapping the distribution of
neurotransmitter receptors rendered accessible yet another
local characteristic of neuronal populations
\citep{zilles2009receptor}.
% allude to our many-regions layer in the methods
In the computational era,
automatic clustering methods are increasingly employed to
regionally differentiation of the cerebral cortex,
which partly turned out to be more fine-grained than
classical cytoarchitectonic mapping
\citep{behrens03, cbp2015review}.
Today,
high-thoughput computing enables 
ultahigh-resolution 3D models of brain anatomy
at macroscopical to near-cellular scale
\citep{amunts2013bigbrain}.
%
As a crucial common point,
all these methodological classes
yielded neuroscientific findings
that were naturally interpreted
non-overlapping, discrete region compartments
as the basic architectonic framework of cortical organization.

@@

a compartmental model of
functional organization in the cerebral cortex can be derived

@@

We assume regions and networks but can estimate 
importance of each voxel.

@@

curse of dimensionality has classically been circumvented
by treating a high-dimensional problem
as a mass-univariate problem.
Problem was postponed to the inference step
this is why we never solved the multiple comparisons problem in fMRI

regularization with sparsity penalties
restrictions to complexity according to neurobiological facts
-> bias-variance tradeoff: negigible increase in bias but large decrease in variance
-> VC: decrease complexitiy capactiy -> better chance for generalization out of sample

multiple comparisons problem has the curse of dimensionality
as its close correspondence in statistical learning
sparsity -> solves the curse at the estimation level

OUTLOOK:
robust high-dimensional inference will be key
to prediction in single individual,
such as for personalized medicine
\citep{gabrieli2015prediction}
